\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants determine interactions with floral visitors

### 2022-06-16 Manuscript/Lanuza_et_al_2021.Rmd
edits

This study investigates the reproductive spectrum of flowering plants and shows that plants have clear positive and negative correlations between reproductive traits that determine interactions with floral visitors at a global scale. Despite the enormous variability of plant reproductive structures, the first two and three axes were able to capture over 50% and 70% of reproductive trait variation, respectively. These three axes were: (1) the flower number - flower size, (2) the pollinator dependence - floral display and (3) the style length - pollinator dependence. Although the explained trait variation that we found in the first two axes is lower than previous studies that consider morphological and physiological traits [@diaz2016; @carmona2021], it is consistent with the largest and most recent study that has characterised plant life strategies with reproductive and plant form and size related traits [@salguero2016]. Interestingly, these different plant reproductive axes were able to partly explain the number of interacting partners of floral visitor guilds. However, they were insufficient to capture their visitation rates which highlights the need to account for other factors to better capture the strength of these mutualistic interactions at a macroecological scale.

Although the negative correlation between flower number and flower size has received substantial attention [worley2000; @sargent2007; @kettle2011], it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. Here, we provide support to this evidence with a great diversity of phylogenetic lineages (170 families and over 1000 species) and show that other reproductive and plant size and form related traits also covariate along this axis. For example, most species with unisexual flowers (14% of all species) were concentrated in the trait space and were associated with species with woddy life forms (i.e., trees and shrubs) with high flower number. This empirical evidence corroborates the idea that approximately 10% of species have separation of sexes [@barrett2002] and that breeding systems are associated with specific life forms [@petit2006]. The pollinator dependence - floral display axis

check number of species of goodwillie and what we offer that is new...

Further, the pollinator dependence - floral display axis provided additional support for the previously described positive association between higher outcrossing rate and larger floral display [@goodwillie2010].

Style length...

 Trait variation on these two axes was associated with the 'fast-slow continuum' in plant [@salguero2016] and animal [@healy2019] life-history strategies, as indicated by the different floral and reproductive biology traits associated with plant height, life form and lifespan. The 'slow' part of this continuum (i.e., tall trees and shrubs) included plant species with many flowers, few ovules, higher pollinator dependence, frequent occurrence of self-incompatibility and more complex breeding systems (e.g., monoecious and dioecious species). In contrast, plant species that employed the 'fast' strategy (i.e., short herbs), had fewer flowers, more ovules, frequent occurrence of self-compatibility and lower pollinator dependence. Further, on the first two axes of trait variation, we found additional support for the previously described positive association between higher outcrossing rate and larger floral display [@goodwillie2010]. The positive correlation between larger floral display and higher pollinator dependence in our dataset further confirmed this trend (Fig. S11).

@@

>>>>

This study investigates the reproductive spectrum of flowering plants and shows that plants have clear positive and negative correlations between reproductive traits that determine interactions with floral visitors at a global scale. Despite the enormous variability of plant reproductive structures, the first two and three axes were able to capture over 50% and 70% of reproductive trait variation, respectively. These three axes were: (1) the flower number - flower size, (2) the pollinator dependence - floral display and (3) the style length - pollinator dependence. Although the explained trait variation that we found in the first two axes is lower than previous studies that consider other morphological and physiological traits [@diaz2016; @carmona2021], our results are consistent with the largest and most recent study that has characterised plant life strategies with reproductive and plant form and size related traits [@salguero2016]. Interestingly, these different plant reproductive axes were able to partly explain the number of interacting partners of floral visitor guilds. However, they were insufficient to capture their visitation rates which highlights the need to account for other factors to better capture the strength of these mutualistic interactions at a macroecological scale.

Our study provides solid evidence for several widely discussed trait associations with the help of a great diversity of phylogenetic lineages (170 families and over 1000 species) within a multitrait context. Although the negative correlation between flower number and flower size has received substantial attention [@worley2000; @sargent2007; @kettle2011], it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. We found support to this trait association and show that other reproductive and plant size and form related traits also covariate along this axis. For example, our study complements previous evidence that links breeding systems and life forms [@petit2006; @friedman2020] by showing that most species with unisexual flowers are concentrated in the trait space and are associated with woody life forms with high flower number. Moreover, we also provide broad support for the previously documented positive association between outcrossing rate and floral display [@goodwillie2010] where species with high pollinator dependence have greater allocation of resources to attraction of floral visitors. Therefore, our results highlight the presence of reproductive trait covariation patterns that are in part mediated by floral visitors in order to ensure plant reproductive success.

Cosas que se pueden mencionar:

- Hablar algo de las main trends

- Factores que puden afectar que no son al número de parejas interactuantes y tasa de visitas: Quizás mencionar en detalle

- Floral rewards

@@

[Trait variation on these two axes was associated with the 'fast-slow continuum' in plant [@salguero2016] and animal [@healy2019] life-history strategies, as indicated by the different floral and reproductive biology traits associated with plant height, life form and lifespan. The 'slow' part of this continuum (i.e., tall trees and shrubs) included plant species with many flowers, few ovules, higher pollinator dependence, frequent occurrence of self-incompatibility and more complex breeding systems (e.g., monoecious and dioecious species). In contrast, plant species that employed the 'fast' strategy (i.e., short herbs), had fewer flowers, more ovules, frequent occurrence of self-compatibility and lower pollinator dependence. Further, on the first two axes of trait variation, we found additional support for the previously described positive association between higher outcrossing rate and larger floral display [@goodwillie2010]. The positive correlation between larger floral display and higher pollinator dependence in our dataset further confirmed this trend (Fig. S11)]()

### 2022-06-16 Manuscript/Lanuza_et_al_2021.tex
edits

This study investigates the reproductive spectrum of flowering plants and shows that plants have clear positive and negative correlations between reproductive traits that determine interactions with floral visitors at a global scale. Despite the enormous variability of plant reproductive structures, the first two and three axes were able to capture over 50\% and 70\% of reproductive trait variation, respectively. These three axes were: (1) the flower number - flower size, (2) the pollinator dependence - floral display and (3) the style length - pollinator dependence. Although the explained trait variation that we found in the first two axes is lower than previous studies that consider morphological and physiological traits (\protect\hyperlink{ref-carmona2021}{Carmona et al., 2021}; \protect\hyperlink{ref-diaz2016}{Díaz et al., 2016}), it is consistent with the largest and most recent study that has characterised plant life strategies with reproductive and plant form and size related traits (\protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}). Interestingly, these different plant reproductive axes were able to partly explain the number of interacting partners of floral visitor guilds. However, they were insufficient to capture their visitation rates which highlights the need to account for other factors to better capture the strength of these mutualistic interactions at a macroecological scale.

Our study provides solid evidence for several widely discussed trait associations with the help of a great diversity of phylogenetic lineages (170 families and over 1000 species) within a multitrait context. Although the negative correlation between flower number and flower size has received substantial attention {[}worley2000; \protect\hyperlink{ref-sargent2007}{Sargent et al.} (\protect\hyperlink{ref-sargent2007}{2007}); \protect\hyperlink{ref-kettle2011}{Kettle et al.} (\protect\hyperlink{ref-kettle2011}{2011}){]}, it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. We found support to this trait association and show that other reproductive and plant size and form related traits also covariate along this axis. For example, our study complements previous evidence that links breeding systems and life forms {[}\protect\hyperlink{ref-petit2006}{Petit \& Hampe} (\protect\hyperlink{ref-petit2006}{2006}); friedman2020{]} by showing that most species with unisexual flowers were concentrated in the trait space with woody life forms with high flower number. Similarly, the second axis of reproductive trait variation also confirmed the previously described positive association between outcrossing rate and floral display (\protect\hyperlink{ref-goodwillie2010}{Goodwillie et al., 2010}).

that is thought to be one of the mechanisms to ensure pollen transfer (\protect\hyperlink{ref-minnaar2019}{Minnaar et al., 2019}).

As expected, species with higher pollinator dependence showed greater allocation of resources to attraction that has been in order to ensure pollen transfer

species with higher pollination dependende had greater allocation to flower number or flower size.

Trait variation on these two axes was associated with the `fast-slow continuum' in plant (\protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}) and animal (\protect\hyperlink{ref-healy2019}{Healy et al., 2019}) life-history strategies, as indicated by the different floral and reproductive biology traits associated with plant height, life form and lifespan. The `slow' part of this continuum (i.e., tall trees and shrubs) included plant species with many flowers, few ovules, higher pollinator dependence, frequent occurrence of self-incompatibility and more complex breeding systems (e.g., monoecious and dioecious species). In contrast, plant species that employed the `fast' strategy (i.e., short herbs), had fewer flowers, more ovules, frequent occurrence of self-compatibility and lower pollinator dependence. Further, on the first two axes of trait variation, we found additional support for the previously described positive association between higher outcrossing rate and larger floral display (\protect\hyperlink{ref-goodwillie2010}{Goodwillie et al., 2010}). The positive correlation between larger floral display and higher pollinator dependence in our dataset further confirmed this trend (Fig. S11).

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-minnaar2019}{}}%
Minnaar, C., Anderson, B., Jager, M. L. de, \& Karron, J. D. (2019). Plant--pollinator interactions along the pathway to paternity. \emph{Annals of Botany}, \emph{123}(2), 225--245.

@@

>>>>

This study investigates the reproductive spectrum of flowering plants and shows that plants have clear positive and negative correlations between reproductive traits that determine interactions with floral visitors at a global scale. Despite the enormous variability of plant reproductive structures, the first two and three axes were able to capture over 50\% and 70\% of reproductive trait variation, respectively. These three axes were: (1) the flower number - flower size, (2) the pollinator dependence - floral display and (3) the style length - pollinator dependence. Although the explained trait variation that we found in the first two axes is lower than previous studies that consider other morphological and physiological traits (\protect\hyperlink{ref-carmona2021}{Carmona et al., 2021}; \protect\hyperlink{ref-diaz2016}{Díaz et al., 2016}), our results are consistent with the largest and most recent study that has characterised plant life strategies with reproductive and plant form and size related traits (\protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}). Interestingly, these different plant reproductive axes were able to partly explain the number of interacting partners of floral visitor guilds. However, they were insufficient to capture their visitation rates which highlights the need to account for other factors to better capture the strength of these mutualistic interactions at a macroecological scale.

Our study provides solid evidence for several widely discussed trait associations with the help of a great diversity of phylogenetic lineages (170 families and over 1000 species) within a multitrait context. Although the negative correlation between flower number and flower size has received substantial attention (\protect\hyperlink{ref-kettle2011}{Kettle et al., 2011}; \protect\hyperlink{ref-sargent2007}{Sargent et al., 2007}; \protect\hyperlink{ref-worley2000}{Worley et al., 2000}), it has only been investigated across a few hundred of Angiosperm species with a two-trait approach. We found support to this trait association and show that other reproductive and plant size and form related traits also covariate along this axis. For example, our study complements previous evidence that links breeding systems and life forms (\protect\hyperlink{ref-friedman2020}{Friedman, 2020}; \protect\hyperlink{ref-petit2006}{Petit \& Hampe, 2006}) by showing that most species with unisexual flowers are concentrated in the trait space and are associated with woody life forms with high flower number. Moreover, we also provide broad support for the previously documented positive association between outcrossing rate and floral display (\protect\hyperlink{ref-goodwillie2010}{Goodwillie et al., 2010}) where species with high pollinator dependence have greater allocation of resources to attraction of floral visitors. Therefore, our results highlight the presence of reproductive trait covariation patterns that are in part mediated by floral visitors in order to ensure plant reproductive success.

Cosas que se pueden mencionar:

\begin{itemize}
\item
  Hablar algo de las main trends
\item
  Factores que puden afectar que no son al número de parejas interactuantes y tasa de visitas: Quizás mencionar en detalle
\item
  Floral rewards
\end{itemize}

@@

\href{}{Trait variation on these two axes was associated with the `fast-slow continuum' in plant (\protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}) and animal (\protect\hyperlink{ref-healy2019}{Healy et al., 2019}) life-history strategies, as indicated by the different floral and reproductive biology traits associated with plant height, life form and lifespan. The `slow' part of this continuum (i.e., tall trees and shrubs) included plant species with many flowers, few ovules, higher pollinator dependence, frequent occurrence of self-incompatibility and more complex breeding systems (e.g., monoecious and dioecious species). In contrast, plant species that employed the `fast' strategy (i.e., short herbs), had fewer flowers, more ovules, frequent occurrence of self-compatibility and lower pollinator dependence. Further, on the first two axes of trait variation, we found additional support for the previously described positive association between higher outcrossing rate and larger floral display (\protect\hyperlink{ref-goodwillie2010}{Goodwillie et al., 2010}). The positive correlation between larger floral display and higher pollinator dependence in our dataset further confirmed this trend (Fig. S11)}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-worley2000}{}}%
Worley, A. C., Baker, A. M., Thompson, J. D., \& Barrett, S. C. (2000). Floral display in narcissus: Variation in flower size and number at the species, population, and individual levels. \emph{International Journal of Plant Sciences}, \emph{161}(1), 69--79.
